### 2021-01-21 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Added schematic illustrating the cloud phase change feedback

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
    and the subsequent \acp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
      Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
    green bars represent the modeled ranges of \acs{ECS} and \acs{TCR},
    respectively, from the different climate model ensembles of the \acf{CMIP}
    with the corresponding \acf{MMM} illustrated with the horizontal black
    lines (the green dot represents output from a single climate model). The
    numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
    \textcite{Meehl2020}.}
  \label{fig:04:historical_ecs_tcr_ranges}
\end{figure}

To illustrate this and put it into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
\unit{K}$ has almost remained unchanged during the last 40 years
\autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
\acs{AR}4 and \acs{AR}5, the corresponding climate model generations from
\acs{CMIP}1, \acs{CMIP}3 and \acs{CMIP}5 more or less agree with this assessed
range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
used in \acs{AR}3 exhibit values of \acs{ECS} above $4.5 \unit{K}$, resulting
in a upper model range of about $0.6 \unit{K}$ higher than the assessed range
\autocite{Mitchell1990, Cubasch2001}. However, these deviations from the
assessed range are insignificant to the ones present in the \acs{CMIP}6
ensemble, which have become evident over the course of 2020. As shown in
\cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper bounds are
more extreme than in any previous climate model generation, resulting in a
\acs{CMIP}6  \ac{ECS} model range of $\range{1.8}{5.6} \unit{K}$ (as of
\TheMonth{}). This increase in range is particularly relevant for the upper
bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
exceeds this upper bound: out of the 42 analyzed \ac{CMIP}6 models, a third (14
models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
several modeling centers independently confirmed high \ac{ECS} values in their
models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble is
given in \cref{sec:04:possible_reasons_high_ecs_cmip6}.

In contrast to \ac{ECS}, \ac{TCR} has only been evaluated since \acs{AR}1 in
1990, in which it is estimated with $2.3 \unit{K}$ using a single climate model
\commentcite{Bretherton1990}{see single green circle in
  \cref{fig:04:historical_ecs_tcr_ranges}}. Over the years, the modeled range
of \ac{TCR} has decreased to about $\range{1.1}{2.6} \unit{K}$ in the
\acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the assessed range of
\ac{TCR}, which was first quoted in \acs{AR}4 with $\range{1.0}{3.0} \unit{K}$
\autocite{Solomon2007}, has been reduced to $\range{1.0}{2.5} \unit{K}$ in
\acs{AR}5 \autocite{Stocker2013}. While the corresponding climate model ranges
from \acs{CMIP}3 and \acs{CMIP}5 agree with these assessed ranges
\autocite{Randall2007, Flato2013}, the upper bound of the \acs{CMIP}6 model
range ($\range{1.3}{3.0} \unit{K}$) is well above the assessed upper bound of
$2.5 \unit{K}$ given by \acs{AR}5. However, due to the correlation of \ac{TCR}
and \ac{ECS} given by \cref{eq:02:tcr_vs_ecs}, this result is not surprising in
light of the large \ac{ECS} range combined with the many high sensitivity
models of the \acs{CMIP}6 ensemble.

@@

Another possible reason for the high climate sensitivity in the \acs{CMIP}6
ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
to cloud feedbacks immediately impact the climate models' climate sensitivity.
The particular feedback mechanism that is thought to be relevant for the high
\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5). As
the climate warms, the predominantly ice clouds over the Southern Ocean in
these climate models become liquid clouds \autocite{McCoy2015}. As a result,
the cloud will get more reflective to the incoming solar radiation since a
cloud consisting of smaller liquid droplets reflects more sunlight than a cloud
consisting of larger ice crystals (for a fixed water content). In addition,
predominantly liquid clouds tend to precipitate less than mixed-phase clouds
(composed of liquid water and ice), leading to more liquid water that stays in
the cloud. Overall, this forms a strong negative shortwave cloud feedback over
the Southern Ocean in climate models from previous \ac{CMIP} generations.

However, it is well-known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water
\autocite{BodasSalcedo2016, McCoy2015, McCoy2016, Tan2016}. For this reason,
the microphysical representation of mixed-phase clouds has been substantially
improved in many \acs{CMIP}6 models \autocite{BodasSalcedo2019, Gettelman2019,
  Mulcahy2020, Williams2020}, leading for example to a better agreement of the
simulated \ac{TOA} shortwave \ac{CRE} with corresponding observations
\autocite{Bock2020}. As a consequence, the strong negative cloud feedback
described above that results from a cloud phase change from ice clouds in the
present-day to liquid clouds in the future is removed (or significantly
reduced) in those \acs{CMIP}6 models that simulate predominantly liquid clouds
over the present-day Southern Ocean. Moreover, this reduction of the negative
shortwave cloud feedback increases the global net cloud feedback
\autocite{BodasSalcedo2019, Tan2016} and eventually the \ac{ECS} of the
affected \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019}.

\begin{figure}[p]
  \centering
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_netcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:a}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_netcre-cmip5_netcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:b}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_swcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:c}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_swcre-cmip5_swcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:d}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_lwcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:e}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_lwcre-cmip5_lwcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:f}
  \end{subfigure}
  \caption{Geographical distributions of the net (a), shortwave (c) and
    longwave (e) \acf{CRE} feedback parameters for the \acs{CMIP}6 \acf{MMM}
    and the corresponding differences to the \acs{CMIP}5 \acs{MMM}, again for
    the net (b), shortwave (d) and longwave (f) components. The \acs{CRE}
    feedback parameters for each grid cell are calculated with a Gregory
    regression of the grid cell's \acs{CRE} versus the change in the
    \acl{GSAT} (see \cref{subsec:02:cloud_feedback_parameters}). Updated from
    \textcite{Bock2020}.}
  \label{fig:04:cloud_feedback_parameters}
\end{figure}

To further confirm this hypothesis, we show the global distributions of
relevant cloud feedback parameters (expressed by the net, shortwave and
longwave \ac{CRE} feedback parameters; see
\cref{subsec:02:cloud_feedback_parameters}) for the \acs{CMIP}6 \ac{MMM} and
its difference to the \acs{CMIP}5 \ac{MMM} in
\cref{fig:04:cloud_feedback_parameters}. First of all, the comparison of
\cref{fig:04:cloud_feedback_parameters:a,fig:04:cloud_feedback_parameters:c,%
  fig:04:cloud_feedback_parameters:e} demonstrates that the shortwave component
dominates the net \ac{CRE} feedback parameter over large swaths of the globe
in the \acs{CMIP}6 \ac{MMM}. The sign change at approximately $60
\unit{\degree S}$ in the shortwave \ac{CRE} feedback parameter (see
\cref{fig:04:cloud_feedback_parameters:b}) shows where the climate models are
switching from simulating almost purely liquid clouds (north of $60
\unit{\degree S}$) to clouds with an increasing ice component (south of $60
\unit{\degree S}$) in their pre-industrial and present-day experiments. The
increasing fraction of ice in the simulated mixed-phase clouds when moving
south supports a negative shortwave cloud feedback on warming in these areas.
\Cref{fig:04:cloud_feedback_parameters:b,fig:04:cloud_feedback_parameters:d}
show a higher shortwave and net \ac{CRE} feedback parameter over the Southern
Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent to the
hypothesis presented in the paragraph above, which claims that the reduced
negative cloud phase change feedback over the Southern Ocean in \acs{CMIP}6
leads to an increased net cloud feedback parameter in comparison to older
\ac{CMIP} generations. This result is further supported by
\textcite{Zelinka2020}, who find a larger positive zonal mean net cloud
feedback parameter in \acs{CMIP}6 compared to \acs{CMIP}5 due to a more
positive (reduced negative) low‐level cloud feedback, primarily in the
extratropics. Moreover, \textcite{Zelinka2020} show that this mainly results
from an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds
for the pre-industrial and present‐day periods, leading to the aforementioned
reduction in the negative cloud phase change feedback on warming and
eventually to higher \ac{ECS} values.

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch04_papers_ecs_tcr_assessment/figs/historical_ecs_tcr_ranges.pdf}
  \caption{Assessed values of the \acf{ECS} (blue bars) and the \acf{TCR} (red
    bars) over the years in the Charney report from 1979 \autocite{Charney1979}
    and the subsequent \acfp{AR} of the \acf{IPCC} \autocite{Mitchell1990,
      Kattenberg1996, Albritton2001, Solomon2007, Stocker2013}. Orange and
    green bars represent the modeled ranges of \acs{ECS} and \acs{TCR},
    respectively, from the different climate model ensembles of the \acf{CMIP}
    with the corresponding \acf{MMM} illustrated with the horizontal black
    lines (the green dot represents output from a single climate model). The
    numbers correspond to individual \acs{CMIP}5 and \acs{CMIP}6 models (see
    \cref{tab:04:ecs_tcr_cmip5,tab:04:ecs_tcr_cmip6}). Updated from
    \textcite{Meehl2020}.}
  \label{fig:04:historical_ecs_tcr_ranges}
\end{figure}

To illustrate this and put it into historical context,
\cref{fig:04:historical_ecs_tcr_ranges} shows the assessed \ac{ECS} and
\ac{TCR} ranges over the years from the Charney report \autocite{Charney1979}
and the different \acp{AR} of the \ac{IPCC} in combination with the
corresponding modeled ranges from the different \ac{CMIP} generations. Since
the Charney report in 1979, the assessed range of \ac{ECS} of $\range{1.5}{4.5}
\unit{K}$ has almost remained unchanged during the last 40 years
\autocite{Charney1979, Mitchell1990, Kattenberg1996, Albritton2001,
  Stocker2013} with the exception of \acs{AR}4, where the lower bound was
temporarily increased to $2.0 \unit{K}$ \autocite{Solomon2007}. For \acs{AR}2,
\acs{AR}4 and \acs{AR}5, the corresponding climate model generations from
\acs{CMIP}1, \acs{CMIP}3 and \acs{CMIP}5 more or less agree with this assessed
range \autocite{Kattenberg1996, Randall2007, Flato2013}. In contrast to that,
some of the early climate models used in \acs{AR}1 and some \acs{CMIP}2 models
used in \acs{AR}3 exhibit values of \acs{ECS} well above $4.5 \unit{K}$,
resulting in a upper model range of about $0.6 \unit{K}$ higher than the
assessed range \autocite{Mitchell1990, Cubasch2001}. However, these deviations
from the assessed range are insignificant to the ones present in the
\acs{CMIP}6 ensemble, which have become evident over the course of 2020. As
shown in \cref{fig:04:historical_ecs_tcr_ranges}, both the lower and upper
bound are more extreme than in any previous climate model generation, resulting
in a \acs{CMIP}6  \ac{ECS} model range of $\range{1.8}{5.6} \unit{K}$ (as of
\TheMonth{}). This increase in range is particularly relevant for the upper
bound, which is about $1.1 \unit{K}$ higher than the assessed upper bound of
$4.5 \unit{K}$. Moreover, it is not only a small fraction of models that
exceeds this upper bound: out of the 42 analyzed \ac{CMIP}6 models, a third (14
models) has an \ac{ECS} above $4.5 \unit{K}$. In addition to our analysis,
several modeling centers independently confirmed high \ac{ECS} values in their
models \autocite{Andrews2019, Gettelman2019, Wyser2020}. A list of possible
reasons for the increased climate sensitivity in the \acs{CMIP}6 ensemble
including aerosol-cloud interactions and changes in the shortwave cloud
feedback over the Southern Ocean is given in
\cref{sec:04:possible_reasons_high_ecs_cmip6}.

In contrast to \ac{ECS}, \ac{TCR} has only been evaluated since \acs{AR}1 in
1990, in which it is estimated with $2.3 \unit{K}$ using only a single climate
model \commentcite{Bretherton1990}{see single green circle in
  \cref{fig:04:historical_ecs_tcr_ranges}}. Over the years, the modeled range
of \ac{TCR} has decreased from about $\range{1.3}{3.8} \unit{K}$ in the
\acs{CMIP}1 models \autocite{Kattenberg1996} to about $\range{1.1}{2.6}
\unit{K}$ in the \acs{CMIP}5 ensemble \autocite{Flato2013}. Similarly, the
assessed range of \ac{TCR}, which was first quoted in \acs{AR}4 with
$\range{1.0}{3.0} \unit{K}$ \autocite{Solomon2007}, has been reduced to
$\range{1.0}{2.5} \unit{K}$ in the subsequent \acs{AR}5 \autocite{Stocker2013}.
While the corresponding climate model ranges from \acs{CMIP}3 and \acs{CMIP}5
agree with these assessed ranges \autocite{Randall2007, Flato2013}, the upper
bound of the \acs{CMIP}6 model range ($\range{1.3}{3.0} \unit{K}$) is well
above the assessed upper bound of $2.5 \unit{K}$ given by \acs{AR}5. However,
unlike \ac{ECS}, also the lower bound of the \acs{CMIP}6 model range increased
compared to the \acs{CMIP}5 ensemble, leading to a similar \ac{TCR} model
spread in both climate model generations. Due to the correlation of \ac{TCR}
and \ac{ECS} given by \cref{eq:02:tcr_vs_ecs}, the increase of \ac{TCR} in many
\acs{CMIP}6 models is not surprising in light of the many high \ac{ECS} models
that are present the \acs{CMIP}6 ensemble.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch04_papers_ecs_tcr_assessment/figs/cloud_phase_change_feedback.pdf}
  \caption{Schematic illustration of the strong negative shortwave cloud phase
    change feedback over the Southern Ocean, which is present in earlier
    \acs{CMIP} model generations like \acs{CMIP}5 (green arrows). Due to
    substantial improvements in the microphysical representation of mixed-phase
    clouds in many \acs{CMIP}6 models, the predominantly ice clouds over the
    Southern Ocean in the present-day climate of previous climate model
    generations have been replaced with clouds that predominantly consist of
    (supercooled) liquid clouds \autocite{BodasSalcedo2019, Gettelman2019,
      Mulcahy2020, Williams2020}. Thus, in these models a cloud phase change
    due to global warming is no longer possible, which removes (or
    significantly reduces) the strong negative cloud phase change feedback
    (orange arrows). This leads to an increase in the global net cloud
    feedback \autocite{BodasSalcedo2019, Tan2016} and the \acf{ECS} in the
    corresponding \acs{CMIP}6 models \autocite{Andrews2019, Gettelman2019},
    which is a possible explanation for the high \acs{ECS} values found in
    many \acs{CMIP}6 models.}
  \label{fig:04:cloud_phase_change_feedback}
\end{figure}

Another possible reason for the high climate sensitivity in the \acs{CMIP}6
ensemble is related to cloud feedbacks \autocite{Bock2020}. As already
discussed in \cref{subsec:02:climate_feedbacks}, uncertainties in cloud
feedbacks are a major source of uncertainty of \ac{ECS} in modern \acp{ESM}
\autocite{Boucher2013, Flato2013}. Thus, changes in processes that are related
to cloud feedbacks immediately impact the climate models' climate sensitivity.
The particular feedback mechanism that is thought to be relevant for the high
\ac{ECS} in \acs{CMIP}6 is connected to cloud phase changes over the Southern
Ocean that are present in earlier \ac{CMIP} generations (\eg{} \acs{CMIP}5).
This so-called \emph{cloud phase change feedback} is illustrated in
\cref{fig:04:cloud_phase_change_feedback} (green arrows). As the climate warms,
the predominantly ice clouds over the Southern Ocean in these climate models
become liquid clouds \autocite{McCoy2015}. As a result, the cloud will get more
reflective to the incoming solar radiation since a cloud consisting of smaller
liquid droplets reflects more sunlight than a cloud consisting of larger ice
crystals (for a fixed water content), which has a cooling effect on the climate
system. In addition, predominantly liquid clouds tend to precipitate less than
mixed-phase clouds (composed of liquid water and ice). This leads to a higher
liquid water content of the cloud and a further amplification of the cooling
effect. Overall, this forms a strong negative shortwave cloud feedback over the
Southern Ocean in climate models from previous \ac{CMIP} generations.

However, it is well-known that the low-level mixed-phase clouds over the
Southern Ocean in these earlier climate models are biased towards too large
amounts of ice crystals and too little amounts of supercooled liquid water
\autocite{BodasSalcedo2016, McCoy2015, McCoy2016, Tan2016}. For this reason,
the microphysical representation of mixed-phase clouds has been substantially
improved in many \acs{CMIP}6 models \autocite{BodasSalcedo2019, Gettelman2019,
  Mulcahy2020, Williams2020}, leading for example to a better agreement of the
simulated \ac{TOA} shortwave \ac{CRE} with corresponding observations
\autocite{Bock2020}. As a consequence, the strong negative cloud feedback
described above that results from a cloud phase change from ice clouds in the
present-day to liquid clouds in the future is removed (or significantly
reduced) in those \acs{CMIP}6 models that simulate predominantly liquid clouds
over the present-day Southern Ocean. This is illustrated in
\cref{fig:04:cloud_phase_change_feedback} (orange arrows). Moreover, this
reduction of the negative shortwave cloud phase change feedback increases the
global net cloud feedback \autocite{BodasSalcedo2019, Tan2016} and eventually
the \ac{ECS} of the affected \acs{CMIP}6 models \autocite{Andrews2019,
  Gettelman2019}, offering a possible explanation for the high \ac{ECS} values
in some \acs{CMIP}6 models.

\begin{figure}[!t]
  \centering
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_netcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:a}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_netcre-cmip5_netcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:b}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_swcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:c}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_swcre-cmip5_swcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:d}
  \end{subfigure}
  \\
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/abs_cmip6_lwcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:e}
  \end{subfigure}
  ~
  \begin{subfigure}[t]{\SmallSubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch04_papers_ecs_tcr_assessment/figs/bias_cmip6_lwcre-cmip5_lwcre.pdf}
    \caption{}
    \label{fig:04:cloud_feedback_parameters:f}
  \end{subfigure}
  \caption{Geographical distributions of the net (a), shortwave (c) and
    longwave (e) \acf{CRE} feedback parameters for the \acs{CMIP}6 \acf{MMM}
    and the corresponding differences to the \acs{CMIP}5 \acs{MMM}, again for
    the net (b), shortwave (d) and longwave (f) components. The \acs{CRE}
    feedback parameters for each grid cell are calculated with a Gregory
    regression of the grid cell's \acs{CRE} versus the change in the
    \acl{GSAT} (see \cref{subsec:02:cloud_feedback_parameters}). Updated from
    \textcite{Bock2020}.}
  \label{fig:04:cloud_feedback_parameters}
\end{figure}

To further confirm this hypothesis, we show the global distributions of
relevant cloud feedback parameters (expressed by the net, shortwave and
longwave \ac{CRE} feedback parameters; see
\cref{subsec:02:cloud_feedback_parameters}) for the \acs{CMIP}6 \ac{MMM} and
its difference to the \acs{CMIP}5 \ac{MMM} in
\cref{fig:04:cloud_feedback_parameters}. First of all, the comparison of
\cref{fig:04:cloud_feedback_parameters:a,fig:04:cloud_feedback_parameters:c,%
  fig:04:cloud_feedback_parameters:e} demonstrates that the shortwave component
dominates the net \ac{CRE} feedback parameter over large swaths of the globe
in the \acs{CMIP}6 \ac{MMM}. The sign change at approximately $60
\unit{\degree S}$ in the shortwave \ac{CRE} feedback parameter (see
\cref{fig:04:cloud_feedback_parameters:b}) shows where the climate models are
switching from simulating almost purely liquid clouds (north of $60
\unit{\degree S}$) to clouds with an increasing ice component (south of $60
\unit{\degree S}$) in their pre-industrial and present-day experiments. The
increasing fraction of ice in the simulated mixed-phase clouds when moving
south supports a negative shortwave cloud feedback on warming in these areas.
\Cref{fig:04:cloud_feedback_parameters:b,fig:04:cloud_feedback_parameters:d}
show a higher shortwave and net \ac{CRE} feedback parameter over the Southern
Ocean in \acs{CMIP}6 compared to \acs{CMIP}5. This is consistent to the
hypothesis presented in the paragraph above, which claims that the reduced
negative cloud phase change feedback over the Southern Ocean in \acs{CMIP}6
leads to an increased net cloud feedback parameter in comparison to older
\ac{CMIP} generations. This result is further supported by
\textcite{Zelinka2020}, who find a larger positive zonal mean net cloud
feedback parameter in \acs{CMIP}6 compared to \acs{CMIP}5 due to a more
positive (reduced negative) low‐level cloud feedback, primarily in the
extratropics. In comparison to earlier \acs{CMIP} generations, the
\acs{CMIP}6 models exhibit weaker increases in the extratropical low‐level
cloud cover and the corresponding liquid water content as a result of global
warming. Moreover, \textcite{Zelinka2020} show that this mainly results from
an increase in the liquid condensate fraction in the \acs{CMIP}6 clouds for
the pre-industrial and present‐day periods, leading to the aforementioned
reduction in the negative cloud phase change feedback on warming and
eventually to higher \ac{ECS} values.
